---
title: "Medical Statistics"
subtitle: "Probability and Probability Distribution"
author: "<strong>Boncho Ku, Ph.D. in Statistics</strong>"
institute: "<br/> Korea Institute of Oriental Medicine"
date: "<br> 2021/09/09 "
output:
  xaringan::moon_reader:
    css: [default, metropolis, metropolis-fonts, ../assets/css/footer.css, ../assets/css/metropolis-ak.css]
    lib_dir: "../docs/libs"
    # lib_dir: "libs"
    nature:
      highlightStyle: github
      titleSlideClass: [middle, left]
      highlightLines: true
      countIncrementalSlides: false
editor_options: 
  chunk_output_type: console
---

```{r setup-knitr, include=FALSE}
options(htmltools.dir.version = FALSE)
# knitr::opts_knit$set(root.dir='..')
knitr::opts_chunk$set(eval = TRUE, 
                      echo = FALSE, 
                      cache = FALSE,
                      include = TRUE,
                      collapse = FALSE,
                      message=FALSE,
                      warning=FALSE, 
                      dependson = NULL,
                      engine = "R", # Chunks will always have R code, unless noted
                      error = TRUE,
                      fig.path="Figures/",  # Set the figure options
                      fig.align = "center", 
                      #fig.width = 7,
                      #fig.height = 7, 
                      fig.keep='all', fig.retina=3)

```

```{r setup-library}
library(MASS)
library(reshape2)
library(plyr)
library(tidyverse)
library(lubridate)
library(readxl)
library(tidyselect)
library(tidystats)
library(glue)
library(here)
library(gt)
library(gtsummary)
library(kableExtra)
library(icons)

```


class: inverse, center, middle

# Probability `r icon_style(fontawesome("dice", style = "solid"), fill = "white")`

---


# Random vs. Deterministic

### Deterministic Experiment 

> The outcome can be predicted with certainty beforehand

#### **Example**

> - Chemical/physical experiment (e.g. $\mathrm{2H + O = H_2O}$, $S = 0.5gt^2$)
> - Adding two numbers (e.g. 8 + 3 = 11)


### Random Experiment 

> The outcome is determined by chance and cannot be predicted with certainty beforehand


#### **Example**

> - Tossing/rolling a coin/dice $\rightarrow$ $\{H, T\}$ or $\{1,2,3,4,5,6\}$
> - throwing a dart on a board
> - A LOTTO/Hitting the jackpot


**Random phenomena**: outcomes of the random experiment


???

Experiments can be categorized into two types: deterministic experiment and random experiment. 

A **deterministic experiment** is one whose outcome may be predicted with certainty beforehand. 

Experiments in the field of chemistry and physics are typical examples of the deterministic experiments. 

For example, when 2 hydrogens are combined with an oxygen, it must result water H2O. 

Or when adding two numbers such as 8 and 3, it becomes 11 not 10 or 12. 

Meanwhile, the random experiment is one whose outcome is determined by chance. 

That is, we assume that 





---


# Sample Space 

> #### **All possible outcomes of a random experiment** $\rightarrow$ $S$

#### Example 

> - Tossing a coin: $S = {H, T}$
> - Rolling a dice:  $S$={`r fontawesome("dice-one", style = "solid")`, `r fontawesome("dice-two", style = "solid")`, 
`r fontawesome("dice-three", style = "solid")`, `r fontawesome("dice-four", style = "solid")`, 
`r fontawesome("dice-five", style = "solid")`, `r fontawesome("dice-six", style = "solid")`} = {
1, 2, 3, 4, 5, 6}


### $S$ can be _finite_ or _infinite_

> - #### **Finite sample space**: $S = \{e_1, e_2, \ldots, e_N\}$ 


> - #### **Infinite sample space** <br><br>
> 
>   - Natural numbers: $\{1, 2, \ldots\}$ $\rightarrow$ **countably infinite sample space** $\rightarrow$ 
**_discrete sample space_**
>   - Many numbers within an interval (e.g. $0\leq x \leq 1$) $\rightarrow$ 
**non-countably infinite sample space** $\rightarrow$ **_non-discrete sample space_**


**In terms of statistics, sampling is a observation of an outcome from the random experiment and its result is called as a sample**


???


---

# `r fontawesome::fa("r-project", fill = "white")` Practice: Sample Space

#### Making a function to create a sample space for simple tools of random experiments (coin and dices)

.pull-left[
```{r, echo=TRUE}
sample_space <- function(
  outcome, times, prob = FALSE
) {
  n <- length(outcome)
  temp <- list()
  for (i in 1:times) {
    temp[[i]] <- outcome
  }
  out <- expand.grid(
    temp, 
    KEEP.OUT.ATTRS = FALSE
  )
  names(out) <- paste(
    "X", 1:times, sep = ""
  )
  if (prob) 
    out$probs <- rep(1, n^times)/(n^times)
  return(out)
}

```
]


.pull-right[
#### rolling a dice once

```{r, echo=TRUE, comment=NA}
sample_space(1:6, times = 1) |> unlist()

```


#### Toss a coin 3 times (all sample space)

```{r, echo=TRUE, eval=FALSE}
sample_space(c("H", "T"), 3)
```

$\rightarrow$ **CHECK RESULTS**


Please check Kerns, G. Jay. Introduction to probability and statistics using r. Lulu. com, 2010.

]


---

# Events

### **A collection of outcomes $\rightarrow$ a subset of the sample space**

> If outcomes of the random experiment belong to the event $A$ $\rightarrow$ the event $A$ has **_occurred_**. 

#### A single event of $A$ $\rightarrow$ a **_simple_** or **_elementary event_** (See pp. 4)

#### Special Events (Basic Set Operation)

> - $S$: Sample space itself $\rightarrow$ _sure_ or _certain event_ (any element of $S$ must occur) 
> - $\varnothing$: **Empty set**, _impossible, empty, or null event_
> - A event occurs **either $A$ or $B$ or both** $\rightarrow$ $A\cup B$ (**_union_** of $A$ and $B$)
> - **both $A$ and $B$** $\rightarrow$ $A\cap B$ (**_intersection_** of $A$ and $B$)
> - The event "**not $A$**" $\rightarrow$ $A^c$ or $A'$ (**_complement_** of $A$)
> - The event " $A$ **but not $B$**" $\rightarrow$ $A - B$ or $A \cap B^c$ (**_set difference_**)
> - a bunch of events $A_1, A_2, A_3, \ldots$ are _mutually exclusive_ or _disjoint_ if $A_i\cap A_j = \varnothing$, 
where $A_i \neq A_j$


**To review the set algebra, see Appendix E.1 in `IPSUR.pdf`**

---

# `r fontawesome::fa("r-project", fill = "white")` Practice: Events

.pull-left[
#### all three coins show identical faces

```{r, echo=TRUE, comment=NA}
S <- sample_space(c("H", "T"), times = 3)
subset(S, X1 == X2 & X1 == X3 & X2 == X3)

```

#### show 2 heads when tossing a coin 3 times

```{r, echo=TRUE, comment=NA}
count <- apply(S, 1, function(x) as.numeric(x == "H"))
numH <- apply(count, 2, sum)
S[which(numH == 2), ]

```

]

.pull-right[
#### more than 2 times of `r fontawesome("dice-three", style = "solid")` rolling 3 dices

```{r, echo=TRUE, results='hide'}
S <- sample_space(1:6, 3)
count <- apply(S, 1, function(x) as.numeric(x == 3))
num3 <- apply(count, 2, sum)
S[which(num3 >= 2), ]

```

#### sum of all three dices are more than 16

```{r, echo=TRUE, comment=NA}
subset(S, X1 + X2 + X3 > 16)
```

]

---

# Concept of Probability


#### **A measure of the chance or uncertainty that assigns a number between 0 to 1**

> - Sure or certain that a event will occurs $\rightarrow$ probability is 100% or 1
> - Sure or certain that a event will not occurs $\rightarrow$ probability is 0 
> - Probability is 1/4 $\rightarrow$ a 25% chance will occur and a 75% chance will not occur
> - Odds against occurance are 75% to 25%, 3 to 1, odds = 1/3


#### **Theoretical Probability**

> - Define a mathemetical function on the sample space $\rightarrow$ probability measure
> - $h$ different ways out of a total of $n$ possible ways $\rightarrow$ 
the probability is $h/n$ (all of outcomes are equally likely)


#### **Emperical Probability**

> With $n$ repetition, $h$ times of occurance were observed $\rightarrow$ the probability is $h/n$ 
(when $n \rightarrow \infty$)


???

For example, tossing a coin repeatedly, in exactly the same way each time. 

Keep a record of total trials and the number of heads observed 

$$P(\mathrm{Heads}) \approx \frac{\mathrm{total~\#~of~flips}}{\mathrm{\#~of~observed~head}}$$
The mathematical background of this approach is the Law of Large Numbers. 

Let $S_n$ count the number of times that $A$ occurred in the $n$ experiments 

$$\frac{S_n}{n} \rightarrow P(A),~ n \rightarrow \infty$$

As the reasoning goes, to learn about the probability of an event A we need only repeat the

random experiment to get a reasonable estimate of the probability’s value, and if we are not

satisfied with our estimate then we may simply repeat the experiment more times all the while

confident that with more and more experiments our estimate will stabilize to the true value.

The frequentist approach is good because it is relatively light on assumptions and does not

worry about symmetry or claims of objectivity like the measure-theoretic approach does. It is

perfect for the spinning coin experiment. One drawback to the method is that one can never

know the exact value of a probability, only a long-run approximation. It also does not work

well with experiments that can not be repeated indefinitely, say, the probability that it will rain

today, the chances that you get will get an A in your Statistics class, or the probability that the

world is destroyed by nuclear war.

---

# Axioms of Probability

> Suppose $S$ is a sample space and each event $A$ of the $S$, $P(A) = p$ called the probability of $A$ 
and satisfies the following axioms: 


- **Axiom 1**: For any event $A \subset S$, $P(A) \geq 0$

- **Axiom 2**: For the sure or certain event $S$, $P(S) = 1$

- **Axiom 3**: If the events $A_1, A_2, A_3, \ldots$ are disjoint (mutually exclusive) events, 


$$P(A_1 \cup A_2 \cup \ldots) = P\left(\bigcup_{i=1}^{\infty} A_i \right) = P(A_1) + P(A_2) + \cdots = \sum_{i=1}^{\infty}P(A_i)$$


???

The intuition behind the axioms: first, the probability of an event should never be negative.

And since the sample space contains all possible outcomes, its probability should be one, 

or 100%. The final axiom may look intimidating, but it simply means that for a sequence of

disjoint events (in other words, sets that do not overlap), their total probability (measure) should

equal the sum of its parts. For example, the chance of rolling a 1 or a 2 on a die is the chance of

rolling a 1 plus the chance of rolling a 2.


---

# Properties of Probability

#### **Some important theroems related to probability**

- **Theorem 1**: If $A \subset B$ then $P(A) \leq P(B)$

- **Theorem 2**: For avery event $A$, $0 \leq P(A) \leq 1$

- **Theorem 3**: For $\varnothing$, $P(\varnothing) = 0$

- **Theorem 4**: $P(A^c) = 1 - P(A)$

- **Theorem 5**: **_The general addition rule_**

$$\begin{aligned}
P(A\cup B) &= P(A) + P(B) - P(A\cap B)~~~\mathrm{more~generally,}\\
P\left(\bigcup_{i=1}^{n}A_i\right) &= \sum_{i=1}^{n}P(A_i) - \sum_{i=1}^{n-1}\sum_{j=i+1}^{n}P(A_i\cap A_j) +\cdots + (-1)^{n-1}P\left(\bigcap_{i=1}^{n}A_i\right)\end{aligned}$$

- **Theorem 6**: Let $B_1, B_2, \ldots, B_n$ be disjoint and partitioned in $A$

$$P(A) = P(A\cap B_1) + P(A\cap B_2) + \cdots + P(A\cap B_n)$$

$\rightarrow$ **The Theorem of Total Probability**

???

Theorem 1 _Proof_. If $A\subset B$, it suffices  $B = A \cap (B\cup A^c)$ and $A\cap (B\cap A^c)=\varnothing$, 
therefore, $P(B) = P(A \cap (B\cup A^c)) = P(A) + P(B\cup A^c)\geq P(A)$

Theorem 3 _Proof_. $\varnothing$ = $S^c$


---

# Assignment of Probabilities

#### If $S$ consists of outcomes $a_1, a_2, \ldots, a_n$ that are **equally likely** in the sense ( $a_{i}\cap a_{j} = \varnothing$ ) that each $a_i$ has equal chance of occurance, by Axiom 2 and 3 

$$P(S) = P(\{a_1, a_2, \ldots, a_n\})=P(a_1\cup a_2 \cup \ldots\cup a_n) = \sum_{i=1}^{n}P(a_i) = 1$$

> #### Since $P(a_i)$ is identical for all $i$, $P(a_i) = 1/N$


#### For an event $A \subset S$ with $k$ elements: e.g. $A=\{a_{i1}, a_{i2}, \ldots, a_{ik}\}$

$$\begin{aligned}
P(A) &= P(a_{i1}) + P(a_{i2}) + \cdots + P(a_{ik}) \\
     &= \frac{1}{N} + \frac{1}{N} + \cdots + \frac{1}{N} \\
     &= \frac{k}{N} = \frac{\#(A)}{\#(S)}
\end{aligned}$$

### $\rightarrow$ Equally Likely Model (ELM) 

> - _equivalent to the theoretical definition of the probability_
> - convenient and popular way to analyze the random experiments


---

# Counting Method

#### Calculate the probability of occurance of the event $A$ $\rightarrow$ just merely **COUNTING** the number of possible outcomes of $A$ and $S$

### Multiplication Rule 

#### Suppose the experiment consisting of two successive steps

- $n_1$: The number of distinct ways for the first step
- $n_2$: The number of distinct ways for the second step

> - **All possible way made by two steps** $\rightarrow$ $n_1 n_2$ ways

#### More generally 

> An experiment composed of $k$ successive steps with $n_1, n_2, \ldots, n_k$ ways, respectively, 
then all possible number of outcomes from an experiment is $n_1 n_2 \cdots n_k$


#### **EXAMPLE** 

> - Pizza topping: {pepperoni, sausage, anchovies, olives, green peppers}
> - Total number of available topping: $2\cdot 2\cdot 2\cdot 2\cdot 2 = 2^5 = 32$


---

# Counting Method

### Permutations 

> Counting the number of different **arrangements**

$$_nP_r = n(n-1)\cdots(n-r+1)$$

> **_the number of permutations of $n$ objects taken $r$ at a time_**

#### A bag with $n$ distinguishable balls and draw $r$ balls at random. How many possible sequences might be observed? 

- **with replacement**: $n^r$
- **without replacement**: $n(n-1)(n-2)\cdots(n-r + 1)$

#### **EXAMPLE**

Required to seat 5 men and 4 women in a row so that the women occupy the even places

> - The men: $_5P_5 = 5! = 120$
> - the women: $_4P_4 = 4! = 24$
> - $\therefore$ number of arrangements = $120 \times 24 = 2880$


???

The number of permutations of $n$ elements is $n!$


---

# Counting Method

### Combinations

> - A counting method for selected objects **without regard to order and replacement**
> - _combinations of $n$ things taken $r$ at a time_ $\rightarrow$ $_nC_r$

$$_nC_r = {n \choose r} = \frac{n!}{r!(n-r)!}$$
It can be written 


$$_nC_r = {n \choose r} = \frac{n(n-1)\cdots(n-r+1)}{r!} = \frac{_nP_r}{r!}$$

> - Permutations $\rightarrow$ interesting in order of arrangements (e.g. $\{abc\} \neq \{cba\}$)
> - In combinations, $\{abc\} = \{cba\}$

#### **EXAMPLE**

> From 7 consonant and 5 vowels, how many words can be formed consisting of 4 different consonants 
and 3 vowels? 















