---
title: "Medical Statistics 2^nd^ Semester Take Home Final Exam"
subtitle: "**Solutions**"
output: 
  pdf_document: 
    includes: 
      in_header: header.tex
  keep_tex: true
  dev: "cairo.pdf"
  citation_pacakge: natbib
# bibliography: references.bib       
editor_options: 
  chunk_output_type: console
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
# knitr::opts_knit$set(root.dir = '..')
# library(officedown)
# library(officer)
# require(formatR)
# 
# fp <- fp_par(
#   text.align = "center", 
#   padding.bottom = 20, padding.top = 120, 
#   border.bottom = fp_border())
# 
# ft <- fp_text(shading.color='#EFEFEF', bold = TRUE)


options(max.print = 1000)
knitr::opts_chunk$set(eval = TRUE,
                      echo = FALSE,
                      cache = FALSE,
                      include = TRUE,
                      collapse = FALSE,
                      tidy = TRUE,
                      tidy.opts = list(blank=FALSE, width.cutoff=120),
                      message = FALSE,
                      warning = FALSE,
                      dependson = NULL,
                      engine = "R", # Chunks will always have R code, unless noted
                      error = TRUE,
                      tab.topcaption = TRUE, 
                      fig.path="Figures/",  # Set the figure options
                      fig.align = "center",
                      fig.keep='all', fig.retina=2)
```

<br/> <br/>

**Name**: \_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_

<br/> <br/>

## **Notice**

-   Please DO SOLVE ANSWERS BY YOURSELVES!!
-   You can use materials from other textbooks, lecture notes, and
    websites but you have to cite materials in every answer.
-   Write down your answers to each question in this document file (make
    a space for answers below the question)

## **Questions**

1.  Which one of the following statements is False?

    a.  In (prospective) cohort study, subjects are enrolled or grouped
        on the basis of their exposure, then are followed to document
        occurrence of disease. **TRUE**
    b.  The probability of a type 1 error is the probability that you
        reject the null hypothesis. **TRUE**
    c.  In test for heterogeneity of meta-analysis, if Higgins
        I^2^\<=25%, studies are regarded homogeneous and the fixed
        effect model of meta-analysis can generally be used. **TRUE**
    d.  To use the two-sample t-test, we need to assume that the data
        from both samples are normally distributed and they have the
        same variances. **TRUE**
    e.  Especially when more than 20% of cells have expected
        frequencies > 5, we need to use Fisher's exact test to determine
        if there are associations between two categorical variables.
        **FALSE**

2.  Suppose all students who officially enrolled in Medical Statistics
    are playing a game called "The Prisoners and Warder". Joel, Minh,
    and Roggers played roles as prisoners and they all had been
    sentenced to death (I'm really sorry to give you guys such a
    role!!). And Noel has a role as a warder. Noel has selected one of
    the prisoners randomly to be pardoned. Noel has already received the
    name which one is pardoned from the governor, but she is not allowed
    to tell to them. Roggers asks to Noel: "If Minh is going to be
    pardoned, give me the name of Joel. If Joel is pardoned, then give
    me Minh's name. If I'm the one to be pardoned, just flip a coin to
    decide whether to name Minh or Joel." Noel reckons for a while and
    decides to tell Roggers that Joel to be executed. Roggers is so
    pleased because he believes that his probability of surviving has
    gone up from 1/3 to 1/2, as it is now between him and Minh to be
    pardoned. Roggers secretly whispered to Minh to tell the brand new
    information. When Minh has heard this news, he reasons that the
    chance of Roggers to be pardoned is not changed at 1/3, but he is
    pleased since Minh's own chance has gone up to 2/3. Which prisoner
    is correct? Please give a detailed explanation of your reasoning.

**ANSWER**

> Exactly an identical structure to the "Monty Hall Problem". Let's
> define the events that Joel, Minh, and Roggers become pardoned before
> hearing from the warden are A, B, and C, respectively. Then
> $P(A) = P(B) = P(C) = 1/3$.

3.  Consider a sample of size 2 drawn without replacement from an urn
    containing three balls, numbered 1, 2, and 3. Let $X$ be the number
    on the first ball drawn and $Y$ the larger of the two numbers drawn

    a.  Find the joint discrete distribution of $X$ and $Y$
    b.  Find the marginal distribution of $Y$
    c.  Find $P(X = 1|Y =3)$
    d.  Find the $\mathrm{Cov}(X, Y)$

**ANSWER**

a.  Joint Distribution of $X$ abd $Y$

```{r}
library(tidyverse)
library(flextable)
library(kableExtra)

tab3 <- data.frame(
  `$X=1$` = c("0", "1/6", "1/6", "1/3"), 
  `$X=2$` = c("0", "1/6", "1/6", "1/3"), 
  `$X=3$` = c("0", "0", "1/3", "1/3"), 
  `$f_Y(y)$` = c("0", "1/3", "2/3", "1"), 
  check.names = FALSE
)
rownames(tab3) <- c("$Y=1$","$Y=2$","$Y=3$", "$f_X(x)$")

tab3 %>% 
  rownames_to_column %>% 
  as_tibble %>% 
  rename(` `= rowname) %>% 
  kbl(
    booktabs = TRUE, 
    escape = FALSE, 
    align = "lcccc"
  ) %>% 
  kable_styling(
    position = "center",
    latex_options = c("hold_position"), 
    full_width = FALSE
  )

```

b.  Marginal distribution of $Y$

```{r}
tab4 <- tibble(
  ` ` = "$f_Y(y)$",
  `$Y=1$` = "0", 
  `$Y=2$` = "1/3", 
  `$Y=3$` = "2/3", 
  `$\\sum$` = "1"
)

tab4 %>% 
  kbl(
    booktabs = TRUE, 
    escape = FALSE, 
    align = "lcccc"
  ) %>% 
  kable_styling(
    position = "center",
    latex_options = c("hold_position"), 
    full_width = FALSE
  )
  
  
```

c.  $P(X=1|Y=3)$

$$P(X=1|Y=3) = \frac{P(X=1, X=3)}{P(Y=3)} = \frac{1/6}{2/3} = \frac{1}{4}$$
d.

$$\begin{aligned}
E(XY) &= 2\frac{1}{6} + 3\frac{1}{6} + 4\frac{1}{6} + 6\frac{1}{6} + 9\frac{1}{3} = \frac{11}{2} = 5.5\\
E(X)  &= 1\frac{1}{3} + 2\frac{1}{3} + 3\frac{1}{3} = 2 \\
E(Y)  &= 2\frac{1}{3} + 3\frac{2}{3} = \frac{8}{3}
\end{aligned}$$

$\therefore \mathrm{Cov}(X,Y)=E(XY) - E(X)E(Y) = 11/2 - 16/3 = 1/6$

4.  Solve the following problems:

    a.  Let $X_1$, $X_2$ and $X_3$ be uncorrelated random variables with
        common variance $\sigma^2$. Find the correlation coefficient
        between $X_1 + X_2$ and $X_2 + X_3$
    b.  Let $X_1$ and $X_2$ be uncorrelated random variables. Find the
        correlation coefficient between $X_1 + X_2$ and $X_2 - X_1$ in
        terms of $\mathrm{Var}(X_1)$ and $\mathrm{Var}(X_2)$.
    c.  Let $X_1$, $X_2$, and $X_3$ be independently distributed random
        variables with common mean $\mu$ and common variance $\sigma^2$.
        Find the correlation coefficient between $X_2 - X_1$ and $X_3$
        and $X_1$.

**ANSWER**

a.  $\mathrm{Cor}(Y=X_1+X_2, Z=X_2+X_3)$

$$\begin{aligned}
\rho_{YZ} &= \frac{\mathrm{Cov}(Y,Z)}{\sigma_Y\sigma_Z}=\frac{E[(X_1+X_2)(X_2+X_3)] - E[X_1+X_2]E[X_2+X_3]}{\sqrt{\mathrm{Var}[X_1+X_2]\mathrm{Var}[X_2+X_3]}} \\
&=\frac{E[X_1X_2+X_2^2+X_1X_3+X_2X_3] - \left\{E[X_1]E[X_2] + E[X_1]E[X_3] + [E[X_2]]^2 + E[X_2]E[X_3] \right \}}{\sqrt{2\sigma^2\cdot2\sigma^2}} \\
&=\frac{\mathrm{Cov}(X_1, X_2) + \mathrm{Cov}(X_1, X_3) + \mathrm{Cov}(X_2, X_3) + E[X_2^2] - [E[X_2]]^2}{2\sigma^2} \\
&= \frac{\sigma^2}{2\sigma^2} = \frac{1}{2}
\end{aligned}$$

b.  $\mathrm{Cor}(X_1+X_2, X_2 - X_1)$

$$\begin{aligned}
\rho_{X_1+X2, X_2-X_1} &= \frac{E[(X_1+X_2)(X_2-X_1)]-E[X_1+X_2]E[X_2-X_1]}{\sqrt{\mathrm{Var}(X_1+X_2)\mathrm{Var}(X_2 - X_1)}} \\
&=\frac{E[X_2^2 - X_1^2]-\{E[X_2]^2 - E[X_1]^2\}}{\sqrt{(\mathrm{Var}[X_1] + \mathrm{Var}[X_2])^2}} \\ 
&=\frac{\mathrm{Var}[X_2] - \mathrm{Var}[X_1]}{\mathrm{Var}[X_1] + \mathrm{Var}[X_2]}
\end{aligned}$$

c.  $\mathrm{Cor}(X_2-X_1, X_3-X_1)$

$$\begin{aligned}
\rho_{X_2-X_1, X_3-X_1} &= \frac{E[(X_2-X_1)(X_3-X_1)]-E[X_2-X_1]E[X_3-X_1]}{\sqrt{\mathrm{Var}(X_2-X_1)\mathrm{Var}(X_3 - X_1)}} \\
&= \frac{E[X_2X_3-X_1X_2-X_1X_3+X_1^2] - \{E(X_2)E(X_3)-E(X_1)E(X_2)-E(X_1)E(X_3)+[E(X_1)^2]\}}{\sqrt{2\sigma^2 2\sigma^2}} \\
&= \frac{\mathrm{Cor}(X_2,X_3) - \mathrm{Cor}(X_1,X_2) - \mathrm{Cor}(X_1, X_3) + E(X_1^2) - [E(X_1)]^2}{2\sigma^2} \\
&= \frac{\sigma^2}{2\sigma^2} = \frac{1}{2}
\end{aligned}$$

5.  When you start R with Rstudio, there is an example dataset named
    with `mtcars`. The `mtcars` dataset was extracted from the 1974
    *Motor Trend* US magazine, and comprise fuel consumption and 10
    aspects of automobile design and performance for 32 automobiles
    (1973 - 74 models). The detailed description of variables in
    `mtcars` dataset can be checked by typing `help(mtcars)` in the
    prompt of R console window. Write R scipts and confirm the results
    for the following questions.

    a.  Extract `mpg` and `disp` variables from `mtcars` dataset and
        restore it in an object `x` and `y`, respectively.

    b.  Calculate mean, standard deviation, coefficient of variation,
        minimum, maximum, median, 25^th^ and 75^th^ quantiles, and
        interquartile range of `x` and `y`.

    c.  Make scatterplot of `x` and `y` and interpret in terms of the
        correlation coefficient between `x` and `y`.

    d.  Assume that `x` is the population of a mile per gallon of all
        automobiles of US from 1973 to 1974. Suppose a sample of size 2
        automobiles are drawn from the population with replacement and
        calculate sample mean. Then repeat the same procedure 10,000
        times (Hint: check the function `sample()`).

        -   Make histogram of 10,000 sample means
        -   Calculate the mean and standard deviation of 10,000 sample
            means
        -   Compare the above results to the population in terms of mean
            and standard deviation: is the mean of 10,000 sample mean is
            approximate to the population mean? In what proportion did
            the standard deviation of the sample mean decrease compared
            to the standard deviation of the population?
            

```{r echo=TRUE, comment=NA, tidy=TRUE, tidy.opts=list(width.cutoff=80)}
x <- mtcars$mpg; y <- mtcars$disp # a
# b
summ_vec <- function(x, ...) {
  m <- mean(x, ...); s <- sd(x, ...) # mean and sd
  cv <- s/m; 
  iqr <- IQR(x, ...)
  out <- c(
    mean = m, sd = s, cv = cv, 
    min = min(x, ...)
    , q25 = quantile(x, 0.25, ...)
    , median = median(x, ...)
    , q75 = quantile(x, 0.75, ...)
    , max = max(x, ...)
    , iqr = iqr
  )
  out
}
summ_vec(x); summ_vec(y)

```

```{r echo=TRUE, eval=FALSE, comment=NA, tidy=TRUE, tidy.opts=list(width.cutoff=80)}
# c
plot(
  x, y, 
  type = "n", 
  bty = "n", 
  main = "Scatterplot of mpg and disp in the mtcars dataset", 
  xlab = "Miles per gallon", 
  ylab = "Displacement"
)
points(x, y, pch=16, cex=2)
abline(lm(y~x), lty=1, lwd=3, col="red")
abline(h = mean(y), lty=2, col="darkgray")
abline(v = mean(x), lty=2, col="darkgray")
text(
  25, 350, 
  bquote(
    paste(rho ==.(
      sprintf(
        "%.2f", cor(x, y)
      )
    )
  )), adj=0, cex=1.5, pos=4)



```


```{r fig.width=10, fig.height=7, out.width="14cm", out.height="10cm", fig.align="center"}
knitr::include_graphics("Figures/unnamed-chunk-4-6.pdf")
```


```{r, echo=TRUE, comment=NA, tidy=TRUE, tidy.opts=list(width.cutoff=80), fig.width=10, fig.height=7, out.width="14cm", out.height="10cm", fig.align="center"}
# d-1
vec <- numeric(10000)
set.seed(20211218)
for (i in 1:10000) {
  vec[i] <- sample(x, 2, replace = TRUE) |> mean()
}
hist(vec)


```


```{r, echo=TRUE, comment=NA, tidy=TRUE, tidy.opts=list(width.cutoff=80)}
#d-2
n <- 10000
mean(vec); 
sd(vec)*((n-1)/n)

#d-3
sd(x); 
sd(x)/sqrt(2)
# The standard deviation of 10,000 sample means with the sample size of 2 decreased 
# in the proportion of sqrt(2).

```

         

6.  A total of 144 women of different ethnic backgrounds were included
    in a cross-sectional study of factors related to blood clotting. We
    compared mean platelet levels in the four groups using a one-way
    ANOVA. It was reasonable to assume Normality and constant variance.
    Fill the following ANOVA table.

```{r, echo=FALSE, message=FALSE}

tab1 <- tibble(
  Group = c(
    "Caucasian",
    "Afro-Caribbean",
    "Mediterranean",
    "Other"
  ),
  `N (\\%)` = c(
    "90 (62.5)",
    "21 (14.6)",
    "19 (13.2)",
    "14 (9.7)"
  ),
  `Mean ( $\\times 10^9$ )` = c(
    "268.1",
    "254.3",
    "281.1",
    "273.3"
  ),
  `SD ($\\times 10^9$)` = c(
    "77.08",
    "67.50",
    "71.09",
    "63.42"
  )
)

kbl(tab1, 
    booktabs = TRUE, 
    escape = FALSE, 
    align = "lccc") %>% 
  kable_styling(
    position = "center",
    latex_options = c("hold_position"), 
    full_width = FALSE
  )


```

Fill the following ANOVA table

```{r}
tab2 <- tibble(
  Source = c(
    "Between Group", 
    "Within Group", 
    "Total"
  ), 
  SS = c(
    "7500.0", 
    "840000.0", 
    "847500.0"
  ), 
  DF = c(
    "3", 
    "(1)", 
    " "
  ), 
  MS = c(
    "2500", 
    "(2)", 
    " "
  ), 
  "F-ratio" = c(
    "(3)", 
    " ", 
    " "
  ), 
  "P-value" = c(
    "0.670", 
    " ", " "
  )
)

kbl(tab2, 
    booktabs = TRUE, 
    escape = FALSE, 
    align = "lccccc") %>% 
  kable_styling(
    position = "center",
    latex_options = c("hold_position"), 
    full_width = FALSE
  )
  

```


**ANSWER**


- $df_{\mathrm{B}} = g - 1 = 4 - 1 = 3$
- $df_{\mathrm{W}} = 144-4 = 140$ (1)
- $MSW = SSW/df_{B} = 840000/140 = 6000$ (2)
- $F_0 = MSB/MSW = 2500/6000 = 0.417$ 


7.  Calculate the sample size for the following questions

    a.  Suppose the response rate of the patient population under study
        after treatment is expected to be around 55% (i.e., $\theta$ =
        0.55). At $\alpha=0.05$, the required sample size for achieving
        an 80% power ($\beta = 0.2$) correctly detecting a difference
        between the post-treatment response rate and the reference value
        say, 35% (i.e., $\theta_0 = 0.35$) is $N=$? (Hint: Test for
        Equality, One sample design)

    b.  Suppose a low density lipidproteins (LDLs) is considered of
        clinically meaningful difference. Assuming that the standard
        deviation is 15% (i.e., population variance is 0.15), the
        required sample size of each group to achieve an 80% power
        ($\beta = 0.2$) at $\alpha=0.05$ for correctly detecting such
        difference of $\mu_2 - \mu_1 = 0.07$ change obtained by normal
        approximation as $N_1, N_2 =$? (Hint: Test for Equality, Two
        sample parallel Design)
        

**ANSWER**

a. Large sample test for proportions (test for equality, one sample design)

$$
n = \frac{(z_{\alpha/2} + z_{\beta})^2 \theta(1-\theta)}{(\theta - \theta_0)^2} = 
\frac{0.55(1-0.55)(1.96 + 0.84)^2}{(0.55 - 0.35)^2} = 48.51 \approx 49
$$
b. Comparing two means (two-sample parallel design)


$$\begin{aligned}
n &= \frac{2(z_{\alpha/2} + z_{\beta})^2}{\eta^2}, ~~ \eta = \frac{\mu_2 - \mu_1}{\hat{\sigma}} \\
&= \frac{2(1.96 + 0.84)^2}{(0.07/0.15)^2} = 72
\end{aligned}$$


The total sample size is $72\times 2 = 144$



