---
title: "Medical Statistics 2^nd^ Semester Take Home Final Exam"
subtitle: "**Due Date: Dec 19 2023 (13:00) to Dec 23 2023 (23:59)**"
output: 
  pdf_document: 
    includes:
      in_header: header.tex
  keep_tex: true
  dev: "cairo.pdf"
  citation_pacakge: natbib
# bibliography: references.bib       
editor_options: 
  chunk_output_type: console
  markdown: 
    wrap: 72
---


```{r setup, include=FALSE}
# knitr::opts_knit$set(root.dir = '..')
# library(officedown)
# library(officer)
# require(formatR)
# 
# fp <- fp_par(
#   text.align = "center", 
#   padding.bottom = 20, padding.top = 120, 
#   border.bottom = fp_border())
# 
# ft <- fp_text(shading.color='#EFEFEF', bold = TRUE)


options(max.print = 1000)
knitr::opts_chunk$set(eval = TRUE,
                      echo = FALSE,
                      cache = FALSE,
                      include = TRUE,
                      collapse = FALSE,
                      tidy = TRUE,
                      tidy.opts = list(blank=TRUE, width.cutoff=60),
                      message = FALSE,
                      warning = FALSE,
                      dependson = NULL,
                      engine = "R", # Chunks will always have R code, unless noted
                      error = TRUE,
                      tab.topcaption = TRUE, 
                      fig.path="Figures/",  # Set the figure options
                      fig.align = "center",
                      fig.keep='all', fig.retina=2)
```


```{r}
library(tidyverse)
library(knitr)
library(kableExtra)


```



<br/> <br/>

**Name**: \_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_

<br/> <br/>

## **Notice**

-   Please **DO SOLVE ANSWERS BY YOURSELVES!!**
-   You can use materials from other textbooks, lecture notes, and
    websites but you have to provide proper **CITATIONS**.
-   Write down your answers in the **MS Word Document** and save your word filename like [`student number-name.docx`, i.e. `202015015-boncho-ku.docx`]. it is admittable to submit your answers by converting your word file to pdf. 
-   If you make up your mind to submit your answers, send an E-mail to Dr. Boncho Ku and Dr. Mimi Ko with the attachment of your answer file.


## **Questions**

1.  Which one of the following statements is False?

    a)  The probability of a type I error is the probability that you reject the 
        null hypothesis when it is true. 

    b)  Subjects are enrolled or grouped on the basis of their exposure, 
        then are followed to document occurrence of disease in prospective cohort 
        study.

    c)  Especially when more than 20% of cells have expected frequencies \< 5, 
        we need to use Fisher's exact test to determine if there are associations 
        between two categorical variables.

    d)  To use the two-sample t-test, we need to assume that the data from both 
        samples are normally distributed and they have the same variances.

    e)  In test for heterogeneity of meta-analysis, if Higgins $I^2 > 75 \%$, 
        studies are regarded homogeneous and the fixed effect model of 
        meta-analysis can generally be used.
        
Answer: **e**

\newpage

2. Suppose all students who officially enrolled in Medical Statistics are 
playing a game called "The Prisoners and Warder". Ari, Umar, and Happy played 
roles as prisoners and they all had been sentenced to death (I'm really sorry 
to give you guys such a role!!).
And Joyce has a role as a warder. Joyce has selected one of the prisoners 
randomly to be pardoned. Joyce has already received the name which one is 
pardoned from the governor, but Joyce is not allowed to tell to them. 
Umar asks to Joyce: "If Ari is going to be pardoned, give me the name of Happy. 
If Happy is pardoned, then give me Ari's name. If I'm the one to be pardoned, 
just flip a coin to decide whether to name Ari or Happy." 
Joyce reckons for a while and decides to tell Umar that Happy to be executed. 
Umar is so pleased because he believes that his probability of surviving has 
gone up from 1/3 to 1/2, as it is now between him and Ari to be pardoned. 
Umar secretly whispered to Ari to tell the brand new information. 
When Ari has heard this news, he reasons that the chance of Umar to be pardoned 
is not changed at 1/3, but he is pleased since Ari's own chance has gone up to 2/3. 
Which prisoner is correct? Please give a detailed explanation of your reasoning. 

**ANSWER**

Let's define the events that Umar, Happy, and Ari become pardoned before
hearing from the warder (Joyce) are A, B, and C, respectively. Then
$P(A) = P(B) = P(C) = 1/3$. Let $b$ be the event that the warder tells Umar 
(A) that Happy (B) is to be executed. Using Bayes' theorem, 

$$
P(A|b) = \frac{P(b|A)P(A)}{P(b|A)P(A) + P(b|B)P(B) + P(b|C)P(C)}
$$
$$\begin{aligned}
P(b|A) &= \frac{P(b\cap A)}{P(A)} = \frac{1/2\cdot 1/3}{1/3} = \frac{1}{2} \\
P(b|B) &= \frac{P(b\cap B)}{P(B)} = \frac{0}{1/3} = 0 \\
P(b|C) &= \frac{P(b\cap C)}{P(C)} = \frac{1/3}{1/3} = 1 \\
\end{aligned}$$
Plugging the above to the equation for $P(A|b)$, then 

$$
P(A|b) = \frac{1/2\cdot1/3}{1/2\cdot1/3 + 0\cdot1/3+1\cdot1/3} = \frac{1}{3}
$$
Similarly, 


$$\begin{aligned}
P(C|b) &= \frac{P(b|C)P(C)}{P(b|A)P(A) + P(b|B)P(B) + P(b|C)P(C)} \\
&=\frac{1\cdot1/3}{1/2} = \frac{2}{3}
\end{aligned}$$

Therefore, Joyce did not provide any information on whether Umar to be 
pardoned or not. However, Ari's chance to be pardoned becomes double after 
hearing Happy (B) is not pardoned. 



\newpage

3. Solve following problems. 

   a. Let $X$ be a random variable having binomial distribution with parameters 
      $n=25$ and $p=0.2$. Evaluate $P(X < \mu_x - 2\sigma_X)$. 
      
   b. If $X$ is a random variable with Poisson distribution satisfying 
      $P(X=0) = P(X=1)$, what is $E(X)$?
   
   c. If $X$ is normally distributed mean 2 and variance 1, find $P(|X - 2| < 1)$. 
   
   d. Suppose $X$ has a binomial distribution with parameters $n$ and $p$. For 
      what $p$ is $\mathrm{var}(X)$ maximized if we assumed $n$ is fixed?


**ANSWER**

a. $X \sim \mathrm{binomial}(n=25, p=0.2)$, then $\mu_X=np=5$, 
$\sigma^2_X=np(1-p)=4$. $P(X<5-4=1)=P(X=0)=(0.8)^{25}$ (There was mistypo in the 
problem a. All students will get the proper score.)

b. Since $X \sim \mathrm{Pois}(\lambda)$, then 

$$
f_X(x; \lambda) = \frac{\lambda^x\exp(-\lambda)}{x!}
$$
$P(X=0) = \exp(-\lambda)$, $P(X=1)=\lambda\exp(-\lambda)$, 
$\exp(-\lambda)=\lambda\exp(-\lambda)$, $\therefore~ \lambda=1$ (There was mistypo 
in the problem b. All students will get the proper score.)

c. $X\sim N(2, 1)$, $P(|X-2|<1 = P(1<X<3) = 0.6826$

d. $X ~ \mathrm{binom}(n, p)$, $\sigma^2_X=np(1-p)$, to find the maximum for p, 

$$
\frac{d}{dp}\sigma^2_X = n - 2pn = 0, ~~ p = \frac{1}{2}
$$
When $p = 1/2$, we can obtain the maximum of $\sigma^2_X$. 


\newpage

4. Let $X$ be the maximum usage time (hour) of an A-manufactured cellular phone after 
   a full charge. Assume that $X$ is normally distributed with mean 75 (hours) and 
   variance $\sigma^2$. If a purchaser of such cellular phones requires that 
   at least 90 percent of the cellular phones are avaiable to use exceeding 60 hours, 
   what is the largest value of $\sigma$ can be and still have the purchaser 
   satisfied? 

**ANSWER**

Since $X\sim N(75, \sigma^2)$ and $P(X > 60) \geq 0.9$, Let $\Phi(X=x)$ be the cumulative distribution function of standardized normal distribution. 

$$\begin{aligned}
\Phi\left(Z=\frac{X-75}{\sigma}\right) & >= 0.9\\
\Phi\left(\frac{60 - 75}{\sigma}\right) & >= 0.9\\
1 - \Phi\left(\frac{60 - 75}{\sigma}\right) & >= 0.9\\
\end{aligned}$$

Therefore, 

$$
\frac{60-75}{\sigma} <= -1.282,~~ \sigma=15/1.282=11.70
$$


\newpage


5. Solve the following problems: 

   a. Let $X_1$, $X_2$ and $X_3$ be uncorrelated random variables with common variance $\sigma^2$. 
   Find the correlation coefficient between $X_1 + X_2$ and $X_2 + X_3$
   b. Let $X_1$ and $X_2$ be uncorrelated random variables. Find the correlation coefficient 
   between $X_1 + X_2$ and $X_2 - X_1$ in terms of $\mathrm{Var}(X_1)$ and $\mathrm{Var}(X_2)$. 
   c. Let $X_1$, $X_2$, and $X_3$ be independently distributed random variables with common mean 
   $\mu$ and common variance $\sigma^2$. Find the correlation coefficient between $X_2 - X_1$ and $X_3 - X_1$. 


a.  $\mathrm{Cor}(Y=X_1+X_2, Z=X_2+X_3)$

$$\begin{aligned}
\rho_{YZ} &= \frac{\mathrm{Cov}(Y,Z)}{\sigma_Y\sigma_Z}=\frac{E[(X_1+X_2)(X_2+X_3)] - E[X_1+X_2]E[X_2+X_3]}{\sqrt{\mathrm{Var}[X_1+X_2]\mathrm{Var}[X_2+X_3]}} \\
&=\frac{E[X_1X_2+X_2^2+X_1X_3+X_2X_3] - \left\{E[X_1]E[X_2] + E[X_1]E[X_3] + [E[X_2]]^2 + E[X_2]E[X_3] \right \}}{\sqrt{2\sigma^2\cdot2\sigma^2}} \\
&=\frac{\mathrm{Cov}(X_1, X_2) + \mathrm{Cov}(X_1, X_3) + \mathrm{Cov}(X_2, X_3) + E[X_2^2] - [E[X_2]]^2}{2\sigma^2} \\
&= \frac{\sigma^2}{2\sigma^2} = \frac{1}{2}
\end{aligned}$$

b.  $\mathrm{Cor}(X_1+X_2, X_2 - X_1)$

$$\begin{aligned}
\rho_{X_1+X2, X_2-X_1} &= \frac{E[(X_1+X_2)(X_2-X_1)]-E[X_1+X_2]E[X_2-X_1]}{\sqrt{\mathrm{Var}(X_1+X_2)\mathrm{Var}(X_2 - X_1)}} \\
&=\frac{E[X_2^2 - X_1^2]-\{E[X_2]^2 - E[X_1]^2\}}{\sqrt{(\mathrm{Var}[X_1] + \mathrm{Var}[X_2])^2}} \\ 
&=\frac{\mathrm{Var}[X_2] - \mathrm{Var}[X_1]}{\mathrm{Var}[X_1] + \mathrm{Var}[X_2]}
\end{aligned}$$

c.  $\mathrm{Cor}(X_2-X_1, X_3-X_1)$

$$\begin{aligned}
\rho_{X_2-X_1, X_3-X_1} &= \frac{E[(X_2-X_1)(X_3-X_1)]-E[X_2-X_1]E[X_3-X_1]}{\sqrt{\mathrm{Var}(X_2-X_1)\mathrm{Var}(X_3 - X_1)}} \\
&= \frac{E[X_2X_3-X_1X_2-X_1X_3+X_1^2] - \{E(X_2)E(X_3)-E(X_1)E(X_2)-E(X_1)E(X_3)+[E(X_1)^2]\}}{\sqrt{2\sigma^2 2\sigma^2}} \\
&= \frac{\mathrm{Cor}(X_2,X_3) - \mathrm{Cor}(X_1,X_2) - \mathrm{Cor}(X_1, X_3) + E(X_1^2) - [E(X_1)]^2}{2\sigma^2} \\
&= \frac{\sigma^2}{2\sigma^2} = \frac{1}{2}
\end{aligned}$$


\newpage


6. When you start R with Rstudio, there is an example dataset named with `mtcars`. 
   The `mtcars` dataset was extracted from the 1974 _Motor Trend_ US magazine, 
   and comprise fuel consumption and 10 aspects of automobile design and 
   performance for 32 automobiles (1973 - 74 models). The detailed description 
   of variables in `mtcars` dataset can be checked by typing `help(mtcars)` 
   in the prompt of R console window. Write R scipts and confirm the results 
   for the following questions.

   a. Extract `mpg` and `disp` variables from `mtcars` dataset and restore it 
      in an object `x` and `y`, respectively. 
   b. Calculate mean, standard deviation, coefficient of variation, minimum, 
      maximum, median, 25^th^ and 75^th^ quantiles, and interquartile range of 
      `x` and `y`. 
   c. Make scatterplot of `x` and `y` and interpret in terms of the correlation 
      coefficient between `x` and `y`. 
   d. Assume that `x` is the population of a mile per gallon of all automobiles 
      of US from 1973 to 1974. Suppose a sample of size 2 automobiles are drawn 
      from the population with replacement and calculate sample mean. 
      Then repeat the same procedure 10,000 times (Hint: check the function `sample()`). 
      - Make histogram of 10,000 sample means.
      - Calculate the mean and standard deviation of 10,000 sample means. 
      - Compare the above results to the population in terms of mean and 
      standard deviation: is the mean of 10,000 sample mean is approximate to 
      the population mean? In what proportion did the standard deviation of the 
      sample mean decrease compared to the standard deviation of the 
      population? 


```{r echo=TRUE, comment=NA, tidy=TRUE, tidy.opts=list(width.cutoff=80)}
x <- mtcars$mpg; y <- mtcars$disp # a
# b
summ_vec <- function(x, ...) {
  m <- mean(x, ...); s <- sd(x, ...) # mean and sd
  cv <- s/m; 
  iqr <- IQR(x, ...)
  out <- c(
    mean = m, sd = s, cv = cv, 
    min = min(x, ...)
    , q25 = quantile(x, 0.25, ...)
    , median = median(x, ...)
    , q75 = quantile(x, 0.75, ...)
    , max = max(x, ...)
    , iqr = iqr
  )
  out
}
summ_vec(x); summ_vec(y)

```

```{r echo=TRUE, eval=FALSE, comment=NA, tidy=TRUE, tidy.opts=list(width.cutoff=80)}
# c
plot(
  x, y, 
  type = "n", 
  bty = "n", 
  main = "Scatterplot of mpg and disp in the mtcars dataset", 
  xlab = "Miles per gallon", 
  ylab = "Displacement"
)
points(x, y, pch=16, cex=2)
abline(lm(y~x), lty=1, lwd=3, col="red")
abline(h = mean(y), lty=2, col="darkgray")
abline(v = mean(x), lty=2, col="darkgray")
text(
  25, 350, 
  bquote(
    paste(rho ==.(
      sprintf(
        "%.2f", cor(x, y)
      )
    )
  )), adj=0, cex=1.5, pos=4)



```

```{r fig.width=10, fig.height=7, out.width="14cm", out.height="10cm", fig.align="center"}
knitr::include_graphics("Figures/unnamed-chunk-3-6.pdf")
```


```{r, echo=TRUE, comment=NA, tidy=TRUE, tidy.opts=list(width.cutoff=80), fig.width=10, fig.height=7, out.width="14cm", out.height="10cm", fig.align="center"}
# d-1
vec <- numeric(10000)
set.seed(20211218)
for (i in 1:10000) {
  vec[i] <- sample(x, 2, replace = TRUE) |> mean()
}
hist(vec)


```


```{r, echo=TRUE, comment=NA, tidy=TRUE, tidy.opts=list(width.cutoff=80)}
#d-2
n <- 10000; mean(vec); 
sd(vec)*((n-1)/n)

#d-3
sd(x); sd(x)/sqrt(2)
# The standard deviation of 10,000 sample means with the sample size of 2 decreased 
# in the proportion of sqrt(2).

```

\newpage


7.  A total of 160 men of different ethnic backgrounds were included in
    a cross-sectional study of factors related to blood clotting. We
    compared mean platelet levels in the four groups using a one-way
    ANOVA. It was reasonable to assume Normality and constant variance.

```{r, echo=FALSE, message=FALSE}
tab1 <- tibble(
  Group = c(
    "Caucasian",
    "Afro-Caribbean",
    "Mediterranean",
    "Other"
  ),
  `N (\\%)` = c(
    "100 (62.5)",
    "18 (11.3)",
    "23 (14.4)",
    "19 (11.9)"
  ),
  `Mean($\\times 10^9$)` = c(
    "262.1",
    "245.3",
    "254.1",
    "254.3"
  ),
  `Standard deviation ($\\times 10^9$)` = c(
    "70.08",
    "69.50",
    "69.09",
    "71.42"
  )
)

kbl(tab1,
    booktabs = TRUE,
    escape = FALSE,
    align = "lccc") %>%
  kable_styling(
    position = "center",
    latex_options = c("hold_position"),
    full_width = FALSE
  )
  

```

Fill the following ANOVA table

```{r}
tab2 <- tibble(
  Source = c(
    "Between Group", 
    "Within Group", 
    "Total"
  ), 
  `Sum of square` = c(
    "15369.0", 
    "1552824.0", 
    "1568193.0"
  ), 
  df = c(
    "(1)", 
    "(2)", 
    " "
  ), 
  `Mean square` = c(
    "(3)", 
    " ", 
    " "
  ), 
  "F-ratio" = c(
    "0.515", 
    " ", 
    " "
  ), 
  "P-value" = c(
    "0.325", 
    " ", " "
  )
)

kableExtra::kbl(
  tab2, 
  booktabs = TRUE, 
  escape = FALSE, 
  align = "lccccc"
) %>% 
  kableExtra::kable_styling(
    position = "center",
    latex_options = c("hold_position"), 
    full_width = FALSE
  ) %>% 
  kableExtra::footnote(
    general = "df, degree of freedom", general_title = "Note: ", 
    footnote_as_chunk = TRUE
  )
  
  

```


**ANSWER**

(1) **3** (2) **156** (3) **5123**


\newpage

8.  Calculate the sample size for the following questions (*Hint*. Test for 
equality, two sample parallel design).

<!-- -->

a)  An active-controlled randomized trial proposes to assess the effectiveness of 
    Drug A in reducing pain. A previous study showed that Drug A can reduce pain 
    score by 5 points from baseline to week 24 with a standard deviation ($\sigma$) 
    of 1.5. A clinically important difference of 0.4 as compared to active drug is 
    considered to be acceptable. (Level of significance = 5%, Power = 80%, Type 
    of test =two-sided).

b)  A placebo-controlled randomized trial proposes to assess the effectiveness 
    of Drug A in curing infants suffering from sepsis. A previous study showed 
    that proportion of subjects cured by Drug A is 60% and a clinically 
    important difference of 15% as compared to placebo is acceptable. 
    (Level of significance = 5%, Power = 80%, Type of test =two-sided)


**ANSWER** 

a. $\delta=0.4$,  $\sigma = 1.5$,  $\alpha = 0.05$, $1-\beta=0.8$

$$
n_1 = n_2 = \frac{2(Z_{1-\alpha/2} + Z_\beta)^2\sigma^2}{\delta^2} = 
\frac{2(1.96 + 0.84)^2(1.5)^2}{0.4^2} = 220.75 \approx 221
$$
```{r, echo=FALSE}
delta <- 0.4; sigma = 1.5
alpha = 0.05; beta = 0.8
N <- (2*(qnorm(1-alpha/2) + qnorm(beta))^2 * sigma^2)/(delta^2)
N <- ceiling(N)

```

The estimated sample size in this study is `r N` per group. 


b. $\theta_1 = 0.6$ and $\theta_2 = 0.45$ since $\delta = 0.15$

$$
n = \frac{(z_{1-\alpha/2} + z_{1-\beta})^2[\theta_1(1-\theta_1) + \theta_2(1-\theta_2)]}{(\theta_1 - \theta_2)^2} = 
\frac{(1.96 + 0.84)^2[0.6(1-0.6) + 0.45(1-0.45)]}{(0.15)^2} = 170.06 \approx 171
$$

```{r, echo=FALSE}
theta1 <- 0.6; theta2 <- 0.45
delta <- 0.15
alpha = 0.05; beta <- 0.8
N <- (qnorm(1-alpha/2) + qnorm(beta))^2*(theta1*(1-theta1) + theta2*(1-theta2))/delta^2
N <- ceiling(N)

```

The estimated sample size in this study is `r N` per group (note that ceiling is generally applied when calculating the sample size. 

