---
title: "Medical Statistics 2^nd^ Semester Take Home Final Exam"
subtitle: "**Due Date: Dec 17 2024 (10:00) to Dec 20 2024 (23:59)**"
output: 
  pdf_document: 
    includes:
      in_header: header.tex
  keep_tex: true
  dev: "cairo.pdf"
  citation_pacakge: natbib
# bibliography: references.bib       
editor_options: 
  chunk_output_type: console
  markdown: 
    wrap: 72
---


```{r setup, include=FALSE}
# knitr::opts_knit$set(root.dir = '..')
# library(officedown)
# library(officer)
# require(formatR)
# 
# fp <- fp_par(
#   text.align = "center", 
#   padding.bottom = 20, padding.top = 120, 
#   border.bottom = fp_border())
# 
# ft <- fp_text(shading.color='#EFEFEF', bold = TRUE)


options(max.print = 1000)
knitr::opts_chunk$set(eval = TRUE,
                      echo = FALSE,
                      cache = FALSE,
                      include = TRUE,
                      collapse = FALSE,
                      tidy = TRUE,
                      tidy.opts = list(blank=TRUE, width.cutoff=60),
                      message = FALSE,
                      warning = FALSE,
                      dependson = NULL,
                      engine = "R", # Chunks will always have R code, unless noted
                      error = TRUE,
                      tab.topcaption = TRUE, 
                      fig.path="Figures/",  # Set the figure options
                      fig.align = "center",
                      fig.keep='all', fig.retina=2)
```


```{r}
library(tidyverse)
library(knitr)
library(kableExtra)


```


<br/> <br/>

**Name**: \_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_

<br/> <br/>

## **Notice**

-   Please **DO SOLVE ANSWERS BY YOURSELVES!!**
-   You can use materials from other textbooks, lecture notes, and
    websites but you have to provide proper **CITATIONS**.
-   Please avoid using LLM based chatbots (i.e. chatGPT, Claude, Gemini, and so on) to make your answers.
-   Write down your answers in the **MS Word Document** and save your word filename like [`student number-name.docx`, i.e. `202015015-boncho-ku.docx`]. it is admittable to submit your answers by converting your word file to pdf. 
-   If you make up your mind to submit your answers, send an E-mail to Dr. Boncho Ku and Dr. Mimi Ko with the attachment of your answer file.


## **Questions**

1.  Which one of the following statements is **TRUE**?

    a)  The probability of a type I error is the probability that you fail to 
        reject the null hypothesis when it is false.

    b)  In a prospective cohort study, subjects are randomly assigned to groups based on their exposure, and then the occurrence of disease is followed.

    c)  When more than 20% of cells have expected frequencies < 5, the chi-square test should always be used, regardless of sample size.

    d)  To use the two-sample t-test, it is not necessary for the data to be normally distributed or for the variances to be equal.

    e)  In a test for heterogeneity of meta-analysis, if Higgins $I^2 > 75 \%$, 
        studies are regarded homogeneous and the fixed effect model of 
        meta-analysis can generally be used.

\newpage


2. Solve following problems. 

   a. Let $X$ be a random variable having binomial distribution with parameters 
      $n=25$ and $p=0.2$. Evaluate $P(X,\mu_x - 2\sigma_X)$. 
      
   b. If $X$ is a random variable with satisfying $P(X=0) = P(X=1)$, what is 
      $E(X)$?
   
   c. If $X$ is normally distributed mean 2 and variance 1, find $P(|X - 2| < 1)$. 
   
   d. Suppose $X$ has a binomial distribution with parameters $n$ and $p$. For 
      what $p$ is $\mathrm{var}(X)$ maximized if we assumed $n$ is fixed?

<!-- \vspace{5cm}   -->


3. Let $X$ be the maximum usage time (hour) of an A-manufactured cellular phone after 
   a full charge. Assume that $X$ is normally distributed with mean 70 (hours) and 
   variance $\sigma^2$. If a purchaser of such cellular phones requires that 
   at least 90 percent of the cellular phones are available to use exceeding 60 hours, 
   what is the largest value of $\sigma$ can be and still have the purchaser 
   satisfied? 

<!-- \newpage -->


4. Solve the following problems: 

   a. Let $X_1$, $X_2$ and $X_3$ be uncorrelated random variables with common variance $\sigma^2$. 
   Find the correlation coefficient between $X_1 + X_2$ and $X_2 + X_3$
   b. Let $X_1$ and $X_2$ be uncorrelated random variables. Find the correlation coefficient 
   between $X_1 + X_2$ and $X_2 - X_1$ in terms of $\mathrm{Var}(X_1)$ and $\mathrm{Var}(X_2)$. 
   c. Let $X_1$, $X_2$, and $X_3$ be independently distributed random variables with common mean 
   $\mu$ and common variance $\sigma^2$. Find the correlation coefficient between $X_2 - X_1$ and $X_3 - X_1$. 


5.  There are five urns, and they are numbered 1 to 5. Each urn contains
    10 balls. Urn $i$ has $i$ defective balls and $10-i$ non-defective
    balls, $i=1,2,\ldots,5$. For example, urn 3 has three defective
    balls and seven non-defective balls. Consider the following random
    experiment: First an urn is selected at random, and then a ball is
    selected at random from the selected urn. Suppose that the
    experimenter does not know which urn was selected. Let's ask two
    questions.

    a)  What is the probability that a defective ball will be selected?

    b)  If we have already selected the ball and noted that it is
        defecvive, what is the probability that it came from urn 5?


6. When you start R with Rstudio, there is an example dataset named with `mtcars`. 
   The `mtcars` dataset was extracted from the 1974 _Motor Trend_ US magazine, 
   and comprise fuel consumption and 10 aspects of automobile design and 
   performance for 32 automobiles (1973 - 74 models). The detailed description 
   of variables in `mtcars` dataset can be checked by typing `help(mtcars)` 
   in the prompt of R console window. Write R scipts and confirm the results 
   for the following questions.

   a. Extract `mpg` and `disp` variables from `mtcars` dataset and restore it 
      in an object `x` and `y`, respectively. 
   b. Calculate mean, standard deviation, coefficient of variation, minimum, 
      maximum, median, 25^th^ and 75^th^ quantiles, and interquartile range of 
      `x` and `y`. 
   c. Make scatterplot of `x` and `y` and interpret in terms of the correlation 
      coefficient between `x` and `y`. 
   d. Assume that `x` is the population of a mile per gallon of all automobiles 
      of US from 1973 to 1974. Suppose a sample of size 2 automobiles are drawn 
      from the population with replacement and calculate sample mean. 
      Then repeat the same procedure 10,000 times (Hint: check the function `sample()`). 
      - Make histogram of 10,000 sample means.
      - Calculate the mean and standard deviation of 10,000 sample means. 
      - Compare the above results to the population in terms of mean and 
      standard deviation: is the mean of 10,000 sample mean is approximate to 
      the population mean? In what proportion did the standard deviation of the 
      sample mean decrease compared to the standard deviation of the 
      population? 


7.  The distribution of grades in a large statistics course is as
    follows:

```{r}
library(tidyverse)
library(kableExtra)
library(flextable)

x <- c(0.1, 0.4, 0.3, 0.1, 0.1)
matrix(x, ncol = 5) %>% 
  as.data.frame -> x
names(x) <- LETTERS[c(1:4, 6)]
  
h <- tibble(`Grade:` = c("Probability"))
tibble(h, x) %>% 
  kbl(booktabs = TRUE, 
      escape = FALSE, 
      align = "lccccc") %>% 
  kable_styling(
    position = "center", 
    latex_options = c("hold_position")
  )
  
```

To calculate student grade point averages, grades are expressed in a
numerical scale with A=4, B=3, and so on down to F=0.

a)  Find the expected value.

b)  Describe your strategy to **simulate** choosing students at random
    and recording their grades.

c)  Based on your strategy described in b), perform the simulation with
    a sample size of size 30 and calculate the mean of their 30 grades
    **using R**.

d)  Repeat c) 10,000 times and calculate the average of 10,000 means.

e)  Make a histogram of 10,000 means.

f)  Describe your conclusion based on the results of a) to e).


<!-- \newpage -->

8.  A pharmaceutical company has developed two types of sleeping pills, namely A and B. 
    Seven healthy adults were randomly selected, and the time it took for them 
    to fall asleep after taking each sleeping pill was measured (assume that time 
    to fall asleep follows a normal distribution with unknown variance 
    ($\sigma^2$ is unknown). The data collected for sleeping pill A and sleeping 
    pill B are as follows:
    
    a) Conduct an appropriate hypothesis test to determine if there is a difference in 
       the effectiveness of sleeping pills A and B. Use a significance level of 0.05. 
    
    b) Calculate the 95% confidence interval for the difference in the mean time to fall 
       asleep between sleeping pills A and B.
       
    c) Calcuate the p-value. 


```{r}
slp_tbl <- 
  tibble(
    ID = 1:7,
    A = c(15, 19, 16, 14, 17, 14, 17),
    B = c(13, 17, 18, 15, 16, 16, 19)
  )
  

kbl(slp_tbl, 
    booktabs = TRUE, 
    escape = FALSE, 
    align = "lcc") %>% 
  kable_styling(
    position = "center",
    latex_options = c("hold_position"), 
    full_width = FALSE
  )

```



<!-- \newpage -->

9.  Describe the relationship between Pearson's correlation coefficients
    and the regression coefficient (slope) of univariate regression
    analysis. You're definitely able to refer to the Internet or other
    textbooks but you have to give descriptions or formulas in your own
    words and cite appropriately.



10.  A total of 180 men of different ethnic backgrounds were included in a 
     cross-sectional study investigating factors related to blood clotting. 
     We compared the mean platelet levels across four groups using a one-way 
     ANOVA. It was reasonable to assume normality and constant variance.


```{r, echo=FALSE, message=FALSE}
tab1 <- tibble(
  Group = c(
    "Caucasian",
    "Afro-Caribbean",
    "Mediterranean",
    "Other"
  ),
  `N (\\%)` = c(
    "110 (61.1)",
    "22 (12.2)",
    "28 (15.6)",
    "20 (11.1)"
  ),
  `Mean($\\times 10^9$)` = c(
    "265.0",
    "248.4",
    "259.2",
    "257.8"
  ),
  `SD ($\\times 10^9$)` = c(
    "72.15",
    "68.30",
    "70.42",
    "69.12"
  )
)

kbl(tab1,
    booktabs = TRUE,
    escape = FALSE,
    align = "lccc") %>%
  kable_styling(
    position = "center",
    latex_options = c("hold_position"),
    full_width = FALSE
  )
  

```

Fill the following ANOVA table

```{r}
tab2 <- tibble(
  Source = c(
    "Between Group", 
    "Within Group", 
    "Total"
  ), 
  SS = c(
    "18540.0", 
    "1549120.0", 
    "1567660.0"
  ), 
  DF = c(
    "(1)", 
    "(2)", 
    " "
  ), 
  MS = c(
    "(3)", 
    "(4)", 
    " "
  ), 
  "F-ratio" = c(
    "(5)", 
    " ", 
    " "
  ), 
  "P-value" = c(
    "0.5520", 
    " ", " "
  )
)

kbl(tab2, 
    booktabs = TRUE, 
    escape = FALSE, 
    align = "lccccc") %>% 
  kable_styling(
    position = "center",
    latex_options = c("hold_position"), 
    full_width = FALSE
  )
  

```

\newpage

11.  Calculate the sample size for the following questions (_Hint_: Test for Equality).


a)  An active-controlled randomized trial proposes to assess the effectiveness of 
    Drug A in reducing blood pressure. A previous study showed that Drug A can 
    reduce blood pressure by 10mmHg from baseline to week 12 with a standard 
    deviation ($\sigma$) of 2.0 mmHg. A clinically important difference of 1.0 
    mmHg as compared to active drug is considered to be acceptable 
    (Level of significance = 5%, Power = 80%, Type of test = two-sided). 


b)  A placebo-controlled randomized trial proposes to assess the effectiveness of 
    Drug A in curing patients with acute respiratory distress syndrome. A previous 
    study showed that proportion of subjects cured by Drug A is 55% and a 
    clinically important difference of 10% as compared to placebo is acceptable 
    (Level of significance = 5%, Power = 80%, Type of test =two-sided)



